---
title: "Violin_plot_figures_SN"
author: "Devika Agarwal"
date: 'Last update: `r date()`'
output:
  html_document:
    fig_height: 5
    fig_width: 5
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
      toc_depth: 5
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
#setwd("~/10x_scN_tara_human/Samples_merged/Nigra_merged_CCA")
#setwd("F:/Oxford_postdoc/10x_scN_tara_human/Samples_merged/Nigra_merged_CCA")
#setwd("/Volumes/Seagate Backup Plus Drive/Oxford_postdoc/10x_scN_tara_human/Samples_merged/Nigra_merged_CCA")
library(ggplot2)
library(reshape)
library(data.table)
library(dplyr)
library(tidyr)
rm(list = ls())
theme_set(theme_bw())
```

# Read in TPM_protein coding genes data

```{r cars}
SN_log_data <- fread("../protein_gene_counts/Nigra_10x_CCA_tpm_protein_coding_genes.txt") # read in TPM expression matrix where rows are genes and columns are cells
SN_log_data <- as.data.frame(SN_log_data)
markergenes <- c("GFAP","OLR1","GINS3","TH","SLC6A3","GAD1","GAD2", "RGS5","CSF1R","MOG","MOBP","PALM2","LGALS1","PPM1G","VCAN") #SN cell types

#markergenes <- c("SNCA","PRKN","UCHL1","PINK1","PARK7","LRRK2","ATP13A2","HTRA2","PLA2G6","FBXO7","VPS35","EIF4G1","DNAJC6","SYNJ1","DNAJC13","CHCHD2","VPS13C","GBA","MAPT","GAK","RAB29","BST1","SMPD1","SCARB2","CTSD","GALC","SLC17A5","ATP6V0A1","KAT8","ALAS1","COQ7","MCC1","CTSB") #PD genes

#markergenes <- c("SNCA","PRKN","UCHL1","PINK1","PARK7","LRRK2","ATP13A2","FBXO7","VPS35","EIF4G1","DNAJC6","SYNJ1","CHCHD2","VPS13C","GBA","MAPT","GAK","SMPD1","SCARB2","SLC17A5","ATP6V0A1","CTSB","DNAJC13") #PD genes reduced

#markergenes <- c("IMPACT","NRSN2", "ENO2", "FAIM2", "TUBA4A","TH","SLC6A3","NXPH4","GAD2","ALDH1A1","VIP") # Mouse DaN maturation and subtype genes
```

# Filter TPM expression to include only markers on interest
```{r pressure, echo=FALSE}
SN_red <- SN_log_data %>% filter(V1 %in% markergenes)
SN_red_t <- as.data.frame(t(SN_red))
tempDF <- SN_red_t
tempDF[] <- lapply(SN_red_t, as.character)
colnames(SN_red_t) <- tempDF[1, ]
SN_red_t <- SN_red_t[-1 ,]
tempDF <- NULL
SN_red_t[] <- lapply(SN_red_t, function(x) as.numeric(as.character(x)))
SN_red_t <- log10((SN_red_t)+1)
SN_red_t$cell_id <- rownames(SN_red_t)
```
#read in Nigra metadata and merge with gene expression data for marker genes
```{r}
metadata <- read.table(file="Metadata_SN_VlnP_figure.txt", sep="\t", header=T)
SN_data <- merge(metadata,SN_red_t,by=c("cell_id"))

#SN_data <- merge(metadata,SN_log_data,by.x="cell_id",by.y="V1") # Can use this directly abd by pass previous few line of code to merge metadata and genes expression if use the FetchData function in Seurat to gene expression for cerrtains with their normalisation , as then genes are columns and cells are row
```


# get Gene marker and corresponding Expresison values in long format
```{r}
data_long <- gather(SN_data, Gene_markers, Expression, RGS5:TH, factor_key=TRUE)
data_long_1 <- arrange(data_long, Gene_markers,Expression)
data_long_1$Gene_markers <- ordered(data_long_1$Gene_markers, levels = c("GFAP","OLR1","GINS3","TH","SLC6A3","GAD1","GAD2", "RGS5","CSF1R","MOG","MOBP","PALM2","LGALS1","PPM1G","VCAN"))     # order the gene markers for the order you desire so that they appear sequentially

```

# violin plots for level 2 cell types in wide format
```{r, fig.height=3, fig.width=7}
#tiff(filename = "F:/Oxford_postdoc/10x_scN_tara_human/Samples_merged/Nigra_merged_CCA/SN_mouse_DaN_p90_subtypes_genes_violin_plot.tiff", width = 7, height = 4,res = 300, units="in")
vln_sn <- ggplot(data_long_1, aes(x=level2class_new, y= Expression, group= interaction(Gene_markers,level2class_new), fill=level2class_new ))
vln_sn <- vln_sn + geom_violin(trim=T, scale="width", adjust=0.5)
vln_sn <- vln_sn + coord_flip()
vln_sn <- vln_sn +  scale_fill_manual("level2class_new", values = c("Microglia" = "#7997FF", "DaNs" = "#FF3300","Astrocytes-1"="#00B92A","ODC-1"="#00BFC4","OPC"="#A3A500","Endothelial"="#800080","Astrocytes-2"="#FA62DB","ODC-2"="#D89000","ODC-3"="#FF6A98","GABA"="#f2bf00" )) 
vln_sn <- vln_sn + theme(axis.text.y = element_text( size = 6, angle = 0, hjust = 1, vjust = 0, face = "plain"))
vln_sn <- vln_sn + theme(axis.text.x =  element_text( size = 4))
vln_sn <- vln_sn + theme(axis.title.y=element_blank())
vln_sn <- vln_sn + theme(axis.title.x = element_text(size=6))
vln_sn <- vln_sn + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
vln_sn <- vln_sn + theme(panel.background = element_blank())
vln_sn <- vln_sn + theme(panel.border = element_blank())
vln_sn <- vln_sn + theme(legend.position="none")
vln_sn <- vln_sn + facet_grid(level2class_new ~Gene_markers,scales="free" , space = "free",shrink=T)
vln_sn <- vln_sn + theme(strip.background = element_blank(), strip.text.y = element_blank())
vln_sn <- vln_sn + theme(strip.background = element_blank(), strip.text.x = element_text(angle = 45, vjust = 1.3, hjust=1.0,size=5))
vln_sn
```

# violin plots for the long format , vertical orientation
```{r, fig.height=8, fig.width=7}
#tiff(filename = "~/Desktop/Figures/SN_PD_genes_level2_vertical_violin_plot.tiff", width = 15 , height = 20,res = 300, units="cm")
#vln_sn <- ggplot(data_long_1, aes(x=level2class_new, y=Expression, group= interaction(Gene_markers,level2class_new), fill=level2class_new ))
#vln_sn <- vln_sn + geom_violin(trim=T, scale="width", adjust=0.5)
#vln_sn <- vln_sn + coord_flip()
#vln_sn <- vln_sn +  scale_fill_manual("level2class_new", values = c("Microglia" = "#7997FF", "DaNs" = "#FF3300","Astrocytes-1"="#00B92A","ODC-1"="#00BFC4","OPC"="#A3A500","Endothelial"="#800080","Astrocytes-2"="#FA62DB","ODC-2"="#D89000","ODC-3"="#FF6A98","GABA"="#f2bf00"  )) 
#vln_sn <- vln_sn + theme(axis.text.y = element_text( size = 4, angle = 0, hjust = 1, vjust = 0, face = "plain"))
#vln_sn <- vln_sn + theme(axis.text.x =  element_blank())
#vln_sn <- vln_sn + theme(axis.title.y=element_blank())
#vln_sn <- vln_sn +  scale_y_continuous(position="right")
#vln_sn <- vln_sn + theme(axis.title.x = element_text(size=8))
#vln_sn <- vln_sn + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
#vln_sn <- vln_sn + theme(panel.background = element_blank())
#vln_sn <- vln_sn + theme(panel.border = element_blank())
#vln_sn <- vln_sn + theme(legend.position="none")
#vln_sn <- vln_sn + facet_grid(Gene_markers ~ level2class_new,scales="free" , space = "free",shrink=T, switch="y")
#vln_sn <- vln_sn + theme(strip.background = element_blank(), strip.text.y = element_text(size=6,angle = 180, hjust = 1, vjust = 0))
#vln_sn <- vln_sn + theme(strip.background = element_blank(), strip.text.x = element_text(angle = 45, vjust = 1.0, hjust=1.0,size=6))
#vln_sn
```


# violin plots for : level 1 cell types wide format
```{r, fig.height=4, fig.width=14}
#tiff(filename = "F:/Oxford_postdoc/10x_scN_tara_human/Samples_merged/Nigra_merged_CCA/SN_PD_genes_level1_violin_plot.tiff", width = 14, height = 4,res = 300, units="in")
#vln_sn <- ggplot(data_long_1, aes(x=level1_celltype, y= Expression, group= interaction(Gene_markers,level1_celltype), fill=level1_celltype ))
#vln_sn <- vln_sn + geom_violin(trim=T, scale="width", adjust=0.5)
#vln_sn <- vln_sn + coord_flip()
#vln_sn <- vln_sn +  scale_fill_manual("level1_celltype", values = c("Microglia" = "#7997FF", "DaNs" = "#ED6141","Astrocytes"="#00B92A","ODC"="#FF61C9","OPC"="#A3A500","Endothelial"="#800080","GABA"="#f2bf00")) 
#vln_sn <- vln_sn + theme(axis.text.y = element_text( size = 7, angle = 0, hjust = 1, vjust = 0, face = "plain"))
#vln_sn <- vln_sn + theme(axis.text.x =  element_text( size = 5))
#vln_sn <- vln_sn + theme(axis.title.y=element_blank())
#vln_sn <- vln_sn + theme(axis.title.x = element_text(size=6))
#vln_sn <- vln_sn + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
#vln_sn <- vln_sn + theme(panel.background = element_blank())
#vln_sn <- vln_sn + theme(panel.border = element_blank())
#vln_sn <- vln_sn + theme(legend.position="none")
#vln_sn <- vln_sn + facet_grid(level1_celltype ~Gene_markers,scales="free" , space = "free",shrink=T)
#vln_sn <- vln_sn + theme(strip.background = element_blank(), strip.text.y = element_blank())
#vln_sn <- vln_sn + theme(strip.background = element_blank(), strip.text.x = element_text(angle = 45, vjust = 1.0, hjust=1.0,size=7))
vln_sn
```


# violin plots  level 1 cell types vertical orientation
```{r, fig.height=8, fig.width=7}
#tiff(filename = "~/Desktop/SN_PD_genes_level1_violin_vertical_plot_UMI_SCT_logged.tiff", width = 15, height = 20,res = 300, units="cm")
#vln_sn <- ggplot(data_long_1, aes(x=level1_class, y= Expression, group= interaction(Gene_markers,level1_class), fill=level1_class ))
#vln_sn <- vln_sn + geom_violin(trim=T, scale="area", adjust=0.5)
#vln_sn <- vln_sn + coord_flip()
#vln_sn <- vln_sn +  scale_fill_manual("level1_class", values = c("Microglia" = "#7997FF", "DaNs" = "#ED6141","Astrocytes"="#00B92A","ODC"="#FF61C9","OPC"="#A3A500","Endothelial"= "#800080","GABA"="#f2bf00")) 
#vln_sn <- vln_sn + theme(axis.text.y = element_text( size = 4, angle = 0, hjust = 1, vjust = 0, face = "plain"))
#vln_sn <- vln_sn + theme(axis.text.x =  element_blank())
#vln_sn <- vln_sn + theme(axis.title.y=element_blank())
#vln_sn <- vln_sn +  scale_y_continuous(position="right")
#vln_sn <- vln_sn + theme(axis.title.x = element_text(size=8))
#vln_sn <- vln_sn + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
#vln_sn <- vln_sn + theme(panel.background = element_blank())
#vln_sn <- vln_sn + theme(panel.border = element_blank())
#vln_sn <- vln_sn + theme(legend.position="none")
#vln_sn <- vln_sn + facet_grid(Gene_markers ~ level1_class,scales="free",shrink=T, switch="y")
#vln_sn <- vln_sn + theme(strip.background = element_blank(), strip.text.y = element_text(size=6,angle = 180, hjust = 1, vjust = 0))
#vln_sn <- vln_sn + theme(strip.background = element_blank(), strip.text.x = element_text(angle = 45, vjust = 1.0, hjust=1.0,size=6))
#vln_sn
```