---
title: "Violin_plot_figures_Cortex"
author: "Devika Agarwal"
date: 'Last update: `r date()`'
output:
  html_document:
    fig_height: 5
    fig_width: 5
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
      toc_depth: 5
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
#setwd("F:/Oxford_postdoc/10x_scN_tara_human/Samples_merged/Cortex_Merged_CCA/Figures")
library(ggplot2)
library(reshape)
library(data.table)
library(dplyr)
library(tidyr)
#rm(list = ls())
theme_set(theme_bw())
```

# Read in TPM_protein coding genes data

```{r cars}
SN_log_data <- fread("./Cortex_SN_tpm_umi_notscaled_pdrisk_genes.txt")
SN_log_data <- as.data.frame(SN_log_data)
#SN_log_data <- SN_log_data[,-22]
metadata <- read.table(file= "Cortex_10x_CCA_metadata.txt", sep="\t", header=T)
#markergenes <-c("AQP4","GFAP","SATB2","SLC17A7","GAD1","GAD2","CSF3R","PLXDC2","MOG","MOBP","OLIG1","VCAN")# DISTINGUISING BETWEEN ALL CELL TYPES

#markergenes <- c("CBLN2","LAMP5","CUX2","RORB", "COL5A2", "IL1RAPL2","TPBG","ETV1", "PCP4","HS3ST5","HTR2C","TLE4","SEMA3D","NTNG2","NR4A2") # EXCITATORY NEURONS

#markergenes <-c("VIP","CNR1","RELN","THSD7B","TAC3","TSHZ2", "LAMP5", "SV2C","COL5A2","EYA4","NOS1","PVALB","SULF1", "RYR1","TAC1","SST","STXBP6", "SYNPR")# IHIBITORY NEURONS, dont need CCK


#markergenes <- c("SNCA","PRKN","UCHL1","PINK1","PARK7","LRRK2","ATP13A2","HTRA2","PLA2G6","FBXO7","VPS35","EIF4G1","DNAJC6","SYNJ1","DNAJC13","CHCHD2","VPS13C","GBA","MAPT","GAK","RAB29","BST1","SMPD1","SCARB2","CTSD","GALC","SLC17A5","ATP6V0A1","KAT8","ALAS1","COQ7","MCC1","CTSB") #PD genes

markergenes <- c("SNCA","PRKN","UCHL1","PINK1","PARK7","LRRK2","ATP13A2","FBXO7","VPS35","EIF4G1","DNAJC6","SYNJ1","CHCHD2","VPS13C","GBA","MAPT","GAK","SMPD1","SCARB2","SLC17A5","ATP6V0A1","CTSB","DNAJC13") #PD genes reduced

#markergenes <- c("PSMB2","PPIH","PSMB4","RAB7A","REEP5","ACTB","NONO","GAPDH","VPS29","EMC7","CHMP2A","SNRPD3") # House Keeping genes Viola
```

# Filter scaled to include only markers on interest
```{r pressure, echo=FALSE}
SN_red <- SN_log_data %>% filter(V1 %in% markergenes)
SN_red_t <- as.data.frame(t(SN_red))
tempDF <- SN_red_t
tempDF[] <- lapply(SN_red_t, as.character)
colnames(SN_red_t) <- tempDF[1, ]
SN_red_t <- SN_red_t[-1 ,]
tempDF <- NULL
SN_red_t[] <- lapply(SN_red_t, function(x) as.numeric(as.character(x)))
SN_red_t <- log10((SN_red_t)+1)
SN_red_t$cell_id <- rownames(SN_red_t)


```
#read in Nigra metadata and merge with gene expression data for marker
```{r}
#SN_data <- merge(metadata,SN_red_t,by.x="X", by.y="cell_id")
SN_data <- merge(metadata,SN_red_t,by.x="X", by.y="cell_id",all.x=T,all.y=F) #for neuronal cell types only
SN_data <- merge(metadata,SN_log_data,by.x="X", by.y="X",all.x=T,all.y=F) #for neuronal cell types only
```


# make Violin plot in long format
```{r}
#data_long <- gather(SN_data, Gene_markers, Expression, CSF3R:OLIG1, factor_key=TRUE)# for all cell types
data_long <- gather(SN_data, Gene_markers, Expression, SNCA:DNAJC13, factor_key=TRUE)
#data_long <- gather(SN_data, Gene_markers, Expression, THSD7B:PVALB, factor_key=TRUE)
#data_long <- gather(SN_data, Gene_markers, Expression, NR4A2:PCP4, factor_key=TRUE)
data_long_1 <- arrange(data_long, Gene_markers,Expression)
data_long_1$Gene_markers <- ordered(data_long_1$Gene_markers, levels = c("SNCA","PRKN","UCHL1","PINK1","PARK7","LRRK2","ATP13A2","FBXO7","VPS35","EIF4G1","DNAJC6","SYNJ1","CHCHD2","VPS13C","GBA","MAPT","GAK","SMPD1","SCARB2","SLC17A5","ATP6V0A1","CTSB","DNAJC13"))      
 #data_long_1$Gene_markers <- ordered(data_long_1$Gene_markers, levels = c("CBLN2","LAMP5","CUX2","RORB", "COL5A2", "IL1RAPL2","TPBG","ETV1","PCP4","HS3ST5","HTR2C","TLE4","SEMA3D","NTNG2","NR4A2")) 
```

# violin plots for the long format, All cell types LEVEL2
```{r, fig.height=6, fig.width=9}
tiff(filename = "F:/Oxford_postdoc/10x_scN_tara_human/Samples_merged/Cortex_Merged_CCA/Figures/Cortex_level2_PDgenes.tiff", width = 14, height = 6,res = 300, units="in")
p <- ggplot(data_long_1, aes(x=ClusterNames_0.8_level_2, y= Expression, group= interaction(Gene_markers,ClusterNames_0.8_level_2), fill=ClusterNames_0.8_level_2 ))
p <- p + geom_violin(trim=T, scale="width", adjust=0.5)
p <- p + coord_flip()
p <- p +  scale_fill_manual("ClusterNames_0.8_level_2", values=c("ODC"="#00B0F6","OPC"="#E76BF3", "Ex1"="#6BB100", "Ex2"="#39B600","Ex3"="#00BA38","Ex4"="#00BA38","Ex5a"="#00BD5F","Ex5b"="#00BD5F","Ex6a"="#00BD5F","Ex6b"="#00BD5F","Ex8"="#00BD5F", "In1a"="#CF9400", "Microglia"="#00BFC4","Astrocyte"="#F8766D", "In1b"="#CF9400","In1c"="#CF9400", "In3"= "#CF9400","In4a"="#CF9400", "In4b"="#CF9400","In5"="#E58700", "In6a"="#C99800", "In6b"="#C99800", "In8"="#C99800")) 
p <- p + theme(axis.text.y = element_text( size = 6, angle = 0, hjust = 1, vjust = 0, face = "plain"))
p <- p + theme(axis.title.y=element_blank())
p <- p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p <- p + theme(panel.background = element_blank())
p <- p + theme(panel.border = element_blank())
p <- p + theme(legend.position="none")
p <- p + facet_grid(ClusterNames_0.8_level_2 ~Gene_markers,scales="free_y" , space = "free",shrink=T)
p <- p + theme(strip.background = element_blank(), strip.text.y = element_blank())
p <- p + theme(strip.background = element_blank(), strip.text.x = element_text(angle = 45, vjust = 1.0, hjust=1.0,size=6))
p
```

# violin plots for thewide format, All cell types LEVEL1
```{r, fig.height=2.5, fig.width=7}
#tiff(filename = "F:/Oxford_postdoc/10x_scN_tara_human/Samples_merged/Cortex_Merged_CCA/Figures/Cortex_level1_PDgenes.tiff", width = 14, height = 4,res = 300, units="in")
Cortex <- ggplot(data_long_1, aes(x=ClusterNames_0.8_level1, y= Expression, group= interaction(Gene_markers,ClusterNames_0.8_level1), fill=ClusterNames_0.8_level1 ))
Cortex <- Cortex + geom_violin(trim=T, scale="width", adjust=0.5)
Cortex <- Cortex + coord_flip()
Cortex <- Cortex +  scale_fill_manual("ClusterNames_0.8_level1", values=c("ODC"="#00B0F6","OPC"="#E76BF3", "Ex neurons"="#39B600","In neurons"="#CF9400", "Microglia"="#00BFC4","Astrocyte"="#F8766D")) 
Cortex <- Cortex + theme(axis.text.y = element_text( size = 6, angle = 0, hjust = 1, vjust = 0, face = "plain"))
Cortex <- Cortex + theme(axis.text.x  = element_text( size = 4,  face = "plain"))
Cortex <- Cortex + theme(axis.title.y=element_blank())
Cortex <- Cortex + theme(axis.title.x = element_text(size=6))
Cortex <- Cortex + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
Cortex <- Cortex + theme(panel.background = element_blank())
Cortex <- Cortex + theme(panel.border = element_blank())
Cortex <- Cortex + theme(legend.position="none")
Cortex <- Cortex + facet_grid(ClusterNames_0.8_level1 ~Gene_markers,scales="free" , space = "free",shrink=T)
Cortex <- Cortex + theme(strip.background = element_blank(), strip.text.y = element_blank())
Cortex <- Cortex + theme(strip.background = element_blank(), strip.text.x = element_text(angle = 45, vjust = 1.0, hjust=1.0,size=6))
Cortex
```

# violin plots for the long format, All cell types LEVEL1 (vertical orientation)
```{r, fig.height=6, fig.width=9}
tiff(filename = "~/Desktop/Cortex_PDrisk_genes_level1_violin_vertical_plot_tpm_umi_notscaled_v1.tiff", width = 15, height = 20,res = 300, units="cm")
p <- ggplot(data_long_1, aes(x=ClusterNames_0.8_level1, y= Expression, group= interaction(Gene_markers,ClusterNames_0.8_level1), fill=ClusterNames_0.8_level1))
p <- p + geom_violin(trim=T, scale="width", adjust=0.5)
#p <- p + coord_flip()
p <- p +  scale_fill_manual("ClusterNames_0.8_level1", values=c("ODC"="#00B0F6","OPC"="#E76BF3", "Ex Neurons"="#39B600","In Neurons"="#CF9400", "Microglia"="#00BFC4","Astrocyte"="#F8766D")) 

p <- p + theme(axis.text.y = element_text( size = 4, angle = 0, hjust = 1, vjust = 0, face = "plain"))
p <- p + theme(axis.text.x =  element_blank())
p <- p + theme(axis.title.y=element_blank())
p <- p +  scale_y_continuous(position="right")
p <- p + theme(axis.title.x = element_text(size=8))
p <- p + xlab("Cortex level1 cell types")
p <- p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p <- p + theme(panel.background = element_blank())
p <- p + theme(panel.border = element_blank())
p <- p + theme(legend.position="none")
p <- p + facet_grid(Gene_markers ~ ClusterNames_0.8_level1,scales="free" , space = "free",shrink=T, switch="y")
p <- p + theme(strip.background = element_blank(), strip.text.y = element_text(size=6,angle = 180, hjust = 1, vjust = 0))
p <- p + theme(strip.background = element_blank(), strip.text.x = element_text(angle = 45, vjust = 1.0, hjust=1.0,size=6))
p


```


# Only plot Neuronal subtypes , Excitatory
```{r, fig.height=3, fig.width=7.0}
#tiff(filename = "C:/Users/Devika/Documents/10x_scN_tara_human/Samples_merged/Cortex_merge_CCA/Cortex_excitatory_neuronal_cell_type_markers_violin_plot.tiff", width = 10, height = 4,res = 300, units="in")
Ex <- ggplot(data_long_1, aes(x=ClusterNames_0.8_level_2, y= Expression, group= interaction(Gene_markers,ClusterNames_0.8_level_2), fill=ClusterNames_0.8_level_2 ))
Ex <- Ex + geom_violin(trim=T, scale="width", adjust=0.5)
Ex <- Ex + coord_flip()
Ex <- Ex +  scale_fill_manual("ClusterNames_0.8_level_2", values=c( "Ex1"="#6BB100", "Ex2"="#39B600","Ex3"="#00BA38","Ex4"="#00BA38","Ex5a"="#00BD5F","Ex5b"="#00BD5F","Ex6a"="#00BD5F","Ex6b"="#00BD5F","Ex8"="#00BD5F")) 
Ex <- Ex + theme(axis.text.y = element_text( size = 6, angle = 0, hjust = 1, vjust = 0, face = "plain"))
Ex <- Ex + theme(axis.text.x  = element_text( size = 4,  face = "plain"))
Ex <- Ex + theme(axis.title.y=element_blank())
Ex <- Ex + theme(axis.title.x = element_text(size=6))
Ex <- Ex + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
Ex <- Ex + theme(panel.background = element_blank())
Ex <- Ex + theme(panel.border = element_blank())
Ex <- Ex + theme(legend.position="none")
Ex <- Ex + facet_grid(Layers ~Gene_markers,scales="free" , space = "free",shrink=T)
Ex <- Ex + theme(strip.background = element_blank(),strip.text.y =  element_text(size=6, face="bold", angle=0))
Ex <- Ex + theme(strip.background = element_blank(), strip.text.x = element_text(angle = 45, vjust = 1.0, hjust=1.0,size=6))
Ex
```

# Only Plot Neuronal Subtypes, inhibitory
```{r,fig.height=4,fig.width=7}
#tiff(filename = "C:/Users/Devika/Documents/10x_scN_tara_human/Samples_merged/Cortex_merge_CCA/Cortex_inhibitory_neuronal_cell_type_markers_violin_plot.tiff", width = 10, height = 4,res = 300, units="in")
In <- ggplot(data_long_1, aes(x=ClusterNames_0.8_level_2, y= Expression, group= interaction(Gene_markers,ClusterNames_0.8_level_2), fill=ClusterNames_0.8_level_2 ))
In <- In + geom_violin(trim=T, scale="width", adjust=0.5)
In <- In + coord_flip()
In <- In +  scale_fill_manual("ClusterNames_0.8_level_2", values=c( "In1a"="#CF9400", "In1b"="#CF9400","In1c"="#CF9400", "In3"= "#CF9400","In4a"="#CF9400", "In4b"="#CF9400","In5"="#E58700", "In6a"="#C99800", "In6b"="#C99800", "In8"="#C99800")) 
In <- In + theme(axis.text.y = element_text( size = 6, angle = 0, hjust = 1, vjust = 0, face = "plain"))
In <- In + theme(axis.text.x  = element_text( size = 4,  face = "plain"))
In <- In + theme(axis.title.y=element_blank())
In <- In + theme(axis.title.x = element_text(size=6))
In <- In + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
In <- In + theme(panel.background = element_blank())
In <- In + theme(panel.border = element_blank())
In <- In + theme(legend.position="none")
In <- In + facet_grid(Layers ~Gene_markers,scales="free" , space = "free",shrink=T)
In <- In + theme(strip.background = element_blank(),strip.text.y =  element_text(size=6, face="bold", angle=0))
In <- In + theme(strip.background = element_blank(), strip.text.x = element_text(angle = 45, vjust = 1.0, hjust=1.0,size=5))
In
```
```{r}
library(grid)
library(gridExtra)
library(ggplot2)
library(lattice)
```

```{r}
plot_A <- arrangeGrob(Cortex, top = textGrob("a", x= unit(0, "npc"), y=unit(1,"npc"), just = c("left","top"), gp= gpar(col="black", fontsize=10,fontfamily="Times Roman", fontface="bold")))

plot_B <- arrangeGrob(Ex, top = textGrob("b", x= unit(0, "npc"), y=unit(1,"npc"), just = c("left","top"), gp= gpar(col="black", fontsize=10,fontfamily="Times Roman", fontface="bold")))

plot_C <- arrangeGrob(In, top = textGrob("c", x= unit(0, "npc"), y=unit(1,"npc"), just = c("left","top"), gp= gpar(col="black", fontsize=10,fontfamily="Times Roman", fontface="bold")))

tiff(filename = "F:/Oxford_postdoc/10x_scN_tara_human/Samples_merged/Cortex_Merged_CCA/Figures/Cortex_celltypes_multipanel.tiff", width = 17, height = 20,res = 300, units="cm")
Row_1_2 <- grid.arrange(plot_A, plot_B,plot_C, nrow = 3, heights=c(2/6,2/6,2/6))
#Row_1_2
dev.off()
```

